The MolSSI \qca Project: An open-source platform to compute, organize, and share quantum chemistry data

### 2020-05-15 response.tex

Reviewer: 1

Comments to the Author(s) 
The manuscript documents an ongoing effort to establish a computational
chemistry resource that provides both extensive data and associated tools for 
computation and analysis.  Their open-source and community-driven approach is
commendable, and the manuscript is publishable after the following points are
taken into consideration.

1) The objectives of the project are ambitious and no doubt many previous
attempts have floundered due to problems arising during their development.
Despite this, there is little discussion on the challenges faced in the 
QCArchive project, what possible solutions were considered, and the reasoning
behind the resolutions of these challenges.  For example, can data from
multiple programs be combined into one data set and, if so, how are the various
program options (the defaults of which can vary widely across packages) treated?
When can levels of theory from two different packages be considered equivalent?
(SCF convergence thresholds, algorithms, DFT grids etc.)  How are excited
states handled?

\textbf{Response:} @Comments This seems like a lot to add and the examples provided show that multiple programs can be combined into a dataset. I have added a single line "For example, unique identifiers for given single point computation are molecule, method, basis, program, and keyword arguments create a unique computation." to help offset some of their arguments. I would be curious about how much to work through here in attempting to answer them.

BPP: Data from multiple programs can be combined into one data set. Data is stored/retrieved/displayed based on the unique identifiers given above. Trying to reconcile the 'same' calculation using different software is beyond the scope of the project, and is one area where previous projects may have stumbled. Excited states are not currently handled, but planned.

2) I downloaded the software and, although I was able to run the examples in the 
manuscript, it felt very much like I had been given a few fish and not taught
how to fish.  Perhaps the authors could include some examples of how to access
help topics, and how to obtain listings of valid keywords and data identifiers.
Without these the experience feels like groping around in the dark.

\textbf{Response: } @Comments This is an open problem and something that is being worked on. This question feel like Ben should answer. Ditto to above.
We have some examples here: https://qcarchive.molssi.org/examples/

BPP: Not sure how to phrase this, but the usability aspect is the first area I am tackling, since it confused me as well. I agree that it can be inconsistent and poorly documented.

3) An important aspect of any collection of data such as the MQCAS is data
integrity.  Provenance meta-data are included, which would allow for bulk data
invalidation if a bug were to be discovered, but what about other potential
problems such as SCF convergence to a saddle point?  Figure 1.  indicates a
review process is carried out on contributions to the database, but no details
are given.  I think for a topic which is so central to the project, more
discussion is required.

\textbf{Response:} @Comments I have added the line "Submitted datasets are reviewed by the \mqcas maintainers for integrity and often require comparisons to literature values to ensure consistency before accepting a contribution." We only started a formal process starting to accept contributions, this is something quite tricky.

4) In Section 3.2 four software components are mentioned (page 7, line 25), yet 
five are summarised in the subsequent bullet points.  Furthermore, only four of
of the five are expanded on in the following sub-subsections (QCSchema is not 
covered).

\textbf{Response:} @Comments We may want to simply remove QCSchema or we can add a small blurb about it. Mostly I want to avoid describing it in too much detail as it requires its own paper with many more authors.

5) It would be interesting to know if there are any efforts being made toward
integrating other chemical databases such as ChemSpider and whether or not 
an HTTP API exists or is being considered.

\textbf{Response:} We have had several conversation with PubChem and Citrine about cross referencing data, but no movement on this front has happened as of yet. QCFractal interacts exclusively over standard TCP/IP (HTTP/REST) protocols that can be accessed from any language as discussed briefly in Fig 4. 
@Comment we may want to add a small discussion on REST interfaces, but I didn't see a good place to slot this in.

==================================================================================

Reviewer: 2

Comments to the Author(s) 
The authors provide an overview of a new archive for quantum-chemical data that is accessible for free for users. It offers a large variety of options, interfaces with many quantum-chemical programs and shows huge potential for further extension. This is without doubt a great initiative. The article is well written and easy to follow and has the potential for a large number of citations as soon as more users have become aware of this initiative.

The article can almost be published as is and I would just like to mention four points for a very minor revision.

1) Ultimately, it is up to the editor to decide on the article type, but as this article does not provide an in-depth review of a methodology published in the literature but revolves more around software for data mining, I feel that a more appropriate article type would be “Software Focus”.

\textbf{Response:} Ok?

2) Could the authors comment on their strategies to ensure that existing databases are monitored for updates? For instance, they use reaction databases such as ACONF, S22, etc. and in the past the reference values for such sets have been updated multiple times. In addition to having access to structures and low-level energies, it is also important for the CMS community to keep track of the currently most accurate reference data.

\textbf{Response:} The authors agree which is why the authors attempt to keep up to date on the literature and update the values where available and have also created the open gateway to adding additional values. An example of this is the S22 dataset where currently the original benchmark values as well as the "S22a" and "S22b" have been added to the dataset.

3) Ref. 63 is not the correct reference for the B3LYP functional. 
\textbf{Response:} @TDC, this is the one I typically use. Do you have a better one?

4) It is common to write “def2-SVP” instead of “def2-svp”.
\textbf{Response:} All instance have of "def2-svp" have been changed to "def2-SVP".

==================================================================================

Reviewer: 3

Comments to the Author(s) 
The review by Smith et al. describes the QCArchive project. This is an initiative by The Molecular Sciences Software Institute (MolSSI). QCArchive is an important resource that aims to organize, curate, and share quantum chemistry data for the benefit of the entire computational molecular sciences community. QCArchive consists of multiple fully open sources of SW modules that control specific functionalities. All are accessible via GitHub. I think this is an extremely useful review that gives an introduction about the basic functionality of QCArchive and provides short snippets of code & how to use it. This manuscript is well written and overall as a great fit for WIRES CMS.

My only minor comments would be:
1.	For the benefit of users and readers, it would be nice is to supplement code snippets with links to the entire example/tutorial in the jupyter notebook.

\textbf{Response:} A link an example notebook has been added to the manuscript (https://docs.qcarchive.molssi.org/projects/QCFractal/en/stable/quickstart.html).

2.	Is QCArchive unique? Are there any comparable efforts in the EU or perhaps in China? It would be great to complement conclusions with comparison (if any).

\textbf{Response:} As far as we are aware QCArchive is unique within the space of quantum chemistry, there are a number of initiatives within quantum materials such as NOMAD and Materials Project that were discussed at the beginning of the paper.

### 2020-05-15 main.tex
Draft response and revisions complete. -TDC

\title{The MolSSI \qca Project: An open-source platform to compute, organize, and share quantum chemistry data}
\author{Daniel G. A. Smith}
\email{dgasmith@vt.edu}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\author{Doaa Altarawy}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\alsoaffiliation{
    Department of Computer and Systems Engineering,
    Alexandria University, Alexandria 21544, Egypt}
\author{Lori A. Burns}
\affiliation{
    Center for Computational Molecular Science and Technology,
    School of Chemistry and Biochemistry,
    Georgia Institute of Technology,
    Atlanta, Georgia 30332-0400, USA}
\author{Matthew Welborn}
\author{Levi N. Naden}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\author{Logan Ward}
\affiliation{
    Data Science and Learning Division
    Argonne National Laboratory,
    Lemont, Illinois 60439, USA}
\author{Sam Ellis}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\author{T. Daniel Crawford}
\affiliation{
    Department of Chemistry, Virginia Tech,
    Blacksburg, Virginia 24061, USA}
\alsoaffiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}

@@

Data-driven research is uniquely positioned to advance the computational molecular sciences (CMS), and the current rapid movement of the field towards modern data science techniques only reinforces this stance.
Successful examples of data-based initiatives in inorganic materials science include the Materials Project\cite{Jain2013} with 110,000+ registered users, the NOMAD Laboratory\cite{Draxl2018NOMAD} with approximately 100 million stored computations, and AFlow\cite{Toher2018AFlow} with nearly three million compounds.
Collectively, these and other efforts are enabling new science in their fields by allowing access to unprecedented quantities of data. \cite{natureed2019, Thygesen2016makingthemost} 
Data repositories also naturally enable more reproducible and replicable science. \cite{NAP25303,Widener2019}
Up to this point, however, there has been no similar-scale initiative for quantum chemistry (QC).
%to fill this gap and few organizations who have the resources to launch such an initiative and collect the associated rewards.

@@

The \qcai software components unify to form a platform that can represent computations as structured data, automatically run those computations on a diverse set of physical resources, store computation results, and provide a Python-based front-end to submit and query computations.
In addition to forming a complete platform, the four software components were carefully constructed to be usable outside of the \qca stack to maximize community utility and reusability.
A brief overview of each component can be found below:

\begin{itemize}
    \item \qcsk - A community built key-value/array description of QC (or QC-like) objects such as molecules and input/output quantities from QC programs. \url{https://github.com/MolSSI/QCSchema}
    \item \qcel - Periodic table information, version-controlled physical constants, molecule parsing, testing infrastructure, and \qcsk models. \url{https://github.com/MolSSI/QCElemental}
    \item \qcng - Quantum chemistry and CMS community program executor with \qcsk enforced input/output models. \url{https://github.com/MolSSI/QCEngine}
    \item \qcf - Distributed task scheduler and executor, database store for chemistry results, and organization of results at scale. \url{https://github.com/MolSSI/QCFractal}
    \item \qcptl - Data querying, visualization, organization, and statistical analysis for chemistry-related results and a front-end client for \qcf. \url{https://github.com/MolSSI/QCPortal}
\end{itemize}

@@

\qcf is the central server of the \qca platform, providing distributed computing, data storage, and data querying capabilities.
Fundamentally \qcf is a campaign manager designed for long-duration (years) deployments to passively manage all computation and data storage requirements for an individual researcher or groups of researchers concurrently.
For reference, the \mqcas has currently run for eight months without issue, serves data to over a thousand unique users per month, and evaluates 1--5M tasks per month.
It is estimated that the \mqcas has the projected capacity to scale by at least 15 fold without optimization.

@@

The second stage involves the specification of the level of theory for each TorsionDrive computation. 
To provide a diverse example, a force field (MMFF94~\cite{halgren1996merck}), machine-learned potential (ANI-1x~\cite{smith2018less}), and a quantum chemistry level of theory (B3LYP-D3(BJ)/def2-SVP~\cite{becke1992density,grimme2011effect,weigend2005a}) are chosen.
For each specification, the geometry optimization program, method, basis set (optionally), and gradient evaluation program are also chosen. Finally, calculations are submitted to the \qcf server's queue (Fig.~\ref{fig:torsiondrive2}).

>>>>

\title{The MolSSI \qca Project: An open-source platform to compute, organize, and share quantum chemistry data}
\author{Daniel G. A. Smith}
\email{dgasmith@vt.edu}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\author{Doaa Altarawy}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\alsoaffiliation{
    Department of Computer and Systems Engineering,
    Alexandria University, Alexandria 21544, Egypt}
\author{Lori A. Burns}
\affiliation{
    Center for Computational Molecular Science and Technology,
    School of Chemistry and Biochemistry,
    Georgia Institute of Technology,
    Atlanta, Georgia 30332-0400, USA}
\author{Matthew Welborn}
\author{Levi N. Naden}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\author{Logan Ward}
\affiliation{
    Data Science and Learning Division
    Argonne National Laboratory,
    Lemont, Illinois 60439, USA}
\author{Sam Ellis}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\author{Benjamin P. Pritchard}
\affiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}
\author{T. Daniel Crawford}
\affiliation{
    Department of Chemistry, Virginia Tech,
    Blacksburg, Virginia 24061, USA}
\alsoaffiliation{
    Molecular Sciences Software Institute,
    Blacksburg, Virginia 24060, USA}

@@

Data-driven research is uniquely positioned to advance the computational molecular sciences (CMS), and the current rapid movement of the field towards modern data science techniques only reinforces this stance.
Successful examples of data-based initiatives in inorganic materials science include the Materials Project\cite{Jain2013} with 110,000+ registered users, the NOMAD Laboratory\cite{Draxl2018NOMAD} with approximately 100 million stored computations, and AFlow\cite{Toher2018AFlow} with nearly three million compounds.

Collectively, these and other efforts (e.g., ioChem-BD,\cite{iochem-bd} the Pitt Quantum Repository,\cite{PQR} and
the Computational Chemistry Comparison and Benchmark DataBase\cite{CCCBDB}) are enabling new science in their fields by allowing access to unprecedented quantities of data. \cite{natureed2019, Thygesen2016makingthemost} 
Data repositories also naturally enable more reproducible and replicable science. \cite{NAP25303,Widener2019}
Up to this point, however, there has been no similar-scale initiative for quantum chemistry (QC).
%to fill this gap and few organizations who have the resources to launch such an initiative and collect the associated rewards.

@@

The \qcai software components unify to form a platform that can represent computations as structured data, automatically run those computations on a diverse set of physical resources, store computation results, and provide a Python-based front-end to submit and query computations.
In addition to forming a complete platform, the five software components were carefully constructed to be usable outside of the \qca stack to maximize community utility and reusability.
A brief overview of each component can be found below:

\begin{itemize}
    \item \qcel - Periodic table information, version-controlled physical constants, molecule parsing, testing infrastructure, and \qcsk models. \url{https://github.com/MolSSI/QCElemental}
    \item \qcng - Quantum chemistry and CMS community program executor with \qcsk enforced input/output models. \url{https://github.com/MolSSI/QCEngine}
    \item \qcf - Distributed task scheduler and executor, database store for chemistry results, and organization of results at scale. \url{https://github.com/MolSSI/QCFractal}
    \item \qcptl - Data querying, visualization, organization, and statistical analysis for chemistry-related results and a front-end client for \qcf. \url{https://github.com/MolSSI/QCPortal}
    \item \qcsk - A community built key-value/array description of QC (or QC-like) objects such as molecules and input/output quantities from QC programs. (This important component, which has been developed somewhat separately from the rest of the \qca package, will be decscribed in a later publication.) \url{https://github.com/MolSSI/QCSchema}
\end{itemize}

@@

\qcf is the central server of the \qca platform, providing distributed computing, data storage, and data querying capabilities.
Fundamentally \qcf is a campaign manager designed for long-duration (years) deployments to passively manage all
computation and data storage requirements for an individual researcher or groups of researchers concurrently. (See
the tutorial at \url{https://docs.qcarchive.molssi.org/projects/QCFractal/en/stable/quickstart.html} for an example of the \qcf ecosystem.)
For reference, the \mqcas has currently run for eight months without issue, serves data to over a thousand unique users per month, and evaluates 1--5M tasks per month.
It is estimated that the \mqcas has the projected capacity to scale by at least 15 fold without optimization.

@@

The second stage involves the specification of the level of theory for each
TorsionDrive computation.  To provide a diverse example, a force field
(MMFF94~\cite{halgren1996merck}), machine-learned potential
(ANI-1x~\cite{smith2018less}), and a quantum chemistry level of theory
(B3LYP-D3(BJ)/def2-SVP~\cite{Becke93,Lee88:LYP,weigend2005a})
are chosen.  For each specification, the geometry optimization program,
method, basis set (optionally), and gradient evaluation program are also
chosen. Finally, calculations are submitted to the \qcf server's queue
(Fig.~\ref{fig:torsiondrive2}).
